---
title: "Block Group Population Attributes"
output: html_document
date: "2025-10-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.width = 10)
```

```{r data}
library(tidyverse)
library(sf)
library(leaflet)

bg <- readRDS("blkgrp_data.RDS") # created in block_group_attributes.R
bg_geo <- read_sf("../geography/cbg/esva_2020blkgrp_clipped.geojson")
bg_geo <- st_transform(bg_geo, crs = 4326)

bg_full <- left_join(bg_geo, bg, by = "GEOID")
```

## Population attributes by block group

Potential population attributes to provide additional context

### Population measures {.tabset}

  - Population count: Estimated number of residents
  - Percent White: Percent of residents who identify as white
  - Percent Black: Percent of residents who identify as Black or African-American
  - Percent Hispanic: Percent of residents who identify as Hispanic or Latino
  - Percent under 18: Percent of residents under age 18
  - Percent 65 and over: Percent of residents aged 65 or over
  - Percent below poverty: Percent of residents living in households below the federal poverty threshold
  
#### Population Count

```{r pmap1}
# Total pop
breaks <- c(0,1,500,1000,2000,2500)
legend_labels <- c("None", "500 or less", "501 to 1000", "1001 to 2000", "2001 or more")
sel_range <- c(min(breaks), max(breaks))

pal <- colorBin("Purples", sel_range, bins = breaks, right = TRUE, reverse = FALSE, na.color = "#808080", pretty = FALSE )

pop_total <- leaflet(bg_full) %>%
  addProviderTiles('CartoDB.Positron') %>%
  addPolygons(data = bg_full,
              weight = 0.5,
              color = "#FFFFFF",
              smoothFactor = 0.2,
              fillColor = ~pal(pop_totalE),
              fillOpacity = 0.7,
              popup = paste0("Value: ", bg_full$pop_totalE)) %>%
  addLegend(position = 'bottomright', 
            pal = pal,
            values = sel_range,
            labFormat = function(type, breaks) {
              return(legend_labels)
            },
            title = "Total Population",
            opacity = 0.7)

pop_total
```

#### Percent White

```{r pmap2}
# White
breaks <- c(0,10,20,40,60,80,100)
legend_labels <- c("10% or less", "10% to 20%", "20% to 40%", "40% to 60%", "60% to 80%", "80% to 100%")
sel_range <- c(min(breaks), max(breaks))

pal <- colorBin("Purples", sel_range, bins = breaks, right = TRUE, reverse = FALSE, na.color = "#808080", pretty = FALSE )

pop_white <- leaflet(bg_full) %>%
  addProviderTiles('CartoDB.Positron') %>%
  addPolygons(data = bg_full,
              weight = 0.5,
              color = "#FFFFFF",
              smoothFactor = 0.2,
              fillColor = ~pal(pop_per_white),
              fillOpacity = 0.7,
              popup = paste0("Value: ", round(bg_full$pop_per_white, 1))) %>%
  addLegend(position = 'bottomright', 
            pal = pal,
            values = sel_range,
            labFormat = function(type, breaks) {
              return(legend_labels)
            },
            title = "Percent White",
            opacity = 0.7)

pop_white
```

#### Percent Black

```{r pmap3}
# Black
breaks <- c(0,10,20,40,60,80,100)
legend_labels <- c("10% or less", "10% to 20%", "20% to 40%", "40% to 60%", "60% to 80%", "80% to 100%")
sel_range <- c(min(breaks), max(breaks))

pal <- colorBin("Purples", sel_range, bins = breaks, right = TRUE, reverse = FALSE, na.color = "#808080", pretty = FALSE )

pop_black <- leaflet(bg_full) %>%
  addProviderTiles('CartoDB.Positron') %>%
  addPolygons(data = bg_full,
              weight = 0.5,
              color = "#FFFFFF",
              smoothFactor = 0.2,
              fillColor = ~pal(pop_per_black),
              fillOpacity = 0.7,
              popup = paste0("Value: ", round(bg_full$pop_per_black, 1))) %>%
  addLegend(position = 'bottomright', 
            pal = pal,
            values = sel_range,
            labFormat = function(type, breaks) {
              return(legend_labels)
            },
            title = "Percent Black",
            opacity = 0.7)

pop_black
```

#### Percent Hispanic

```{r pmap4}
# Hispanic
breaks <- c(0,10,20,40,60,80,100)
legend_labels <- c("10% or less", "10% to 20%", "20% to 40%", "40% to 60%", "60% to 80%", "80% to 100%")
sel_range <- c(min(breaks), max(breaks))

pal <- colorBin("Purples", sel_range, bins = breaks, right = TRUE, reverse = FALSE, na.color = "#808080", pretty = FALSE )

pop_hisp <- leaflet(bg_full) %>%
  addProviderTiles('CartoDB.Positron') %>%
  addPolygons(data = bg_full,
              weight = 0.5,
              color = "#FFFFFF",
              smoothFactor = 0.2,
              fillColor = ~pal(pop_per_hisp),
              fillOpacity = 0.7,
              popup = paste0("Value: ", round(bg_full$pop_per_hisp, 1))) %>%
  addLegend(position = 'bottomright', 
            pal = pal,
            values = sel_range,
            labFormat = function(type, breaks) {
              return(legend_labels)
            },
            title = "Percent Hispanic",
            opacity = 0.7)

pop_hisp
```

#### Percent under 18

```{r pmap5}
# Under 18
breaks <- c(0,10,20,40,60,80,100)
legend_labels <- c("10% or less", "10% to 20%", "20% to 40%", "40% to 60%", "60% to 80%", "80% to 100%")
sel_range <- c(min(breaks), max(breaks))

pal <- colorBin("Purples", sel_range, bins = breaks, right = TRUE, reverse = FALSE, na.color = "#808080", pretty = FALSE )

pop_u18 <- leaflet(bg_full) %>%
  addProviderTiles('CartoDB.Positron') %>%
  addPolygons(data = bg_full,
              weight = 0.5,
              color = "#FFFFFF",
              smoothFactor = 0.2,
              fillColor = ~pal(pop_per_under18),
              fillOpacity = 0.7,
              popup = paste0("Value: ", round(bg_full$pop_per_under18, 1))) %>%
  addLegend(position = 'bottomright', 
            pal = pal,
            values = sel_range,
            labFormat = function(type, breaks) {
              return(legend_labels)
            },
            title = "Percent under 18",
            opacity = 0.7)

pop_u18
```

#### Percent 65 and over

```{r pmap6}
# 65 and over
breaks <- c(0,10,20,40,60,80,100)
legend_labels <- c("10% or less", "10% to 20%", "20% to 40%", "40% to 60%", "60% to 80%", "80% to 100%")
sel_range <- c(min(breaks), max(breaks))

pal <- colorBin("Purples", sel_range, bins = breaks, right = TRUE, reverse = FALSE, na.color = "#808080", pretty = FALSE )

pop_65o <- leaflet(bg_full) %>%
  addProviderTiles('CartoDB.Positron') %>%
  addPolygons(data = bg_full,
              weight = 0.5,
              color = "#FFFFFF",
              smoothFactor = 0.2,
              fillColor = ~pal(pop_per_65over),
              fillOpacity = 0.7,
              popup = paste0("Value: ", round(bg_full$pop_per_65over, 1))) %>%
  addLegend(position = 'bottomright', 
            pal = pal,
            values = sel_range,
            labFormat = function(type, breaks) {
              return(legend_labels)
            },
            title = "Percent 65 or over",
            opacity = 0.7)

pop_65o
```

#### Percent below poverty

```{r pmap7}
# poverty
breaks <- c(0,10,20,40,60,80,100)
legend_labels <- c("10% or less", "10% to 20%", "20% to 40%", "40% to 60%", "60% to 80%", "80% to 100%")
sel_range <- c(min(breaks), max(breaks))

pal <- colorBin("Purples", sel_range, bins = breaks, right = TRUE, reverse = FALSE, na.color = "#808080", pretty = FALSE )

pop_pov <- leaflet(bg_full) %>%
  addProviderTiles('CartoDB.Positron') %>%
  addPolygons(data = bg_full,
              weight = 0.5,
              color = "#FFFFFF",
              smoothFactor = 0.2,
              fillColor = ~pal(pop_per_poverty),
              fillOpacity = 0.7,
              popup = paste0("Value: ", round(bg_full$pop_per_poverty, 1))) %>%
  addLegend(position = 'bottomright', 
            pal = pal,
            values = sel_range,
            labFormat = function(type, breaks) {
              return(legend_labels)
            },
            title = "Percent below Poverty",
            opacity = 0.7)

pop_pov
```

### Household measures {.tabset}

  - Housing count: Estimated number of housing units
  - Percent vacant housing: Percent of vacant housing units
  - Median income: Median income of households
  - Percent high income: Percent of households with household incomes $200,000 or more
  - Percent renting: Percent of households who are renting their home
  - Median rent: Median monthly rent paid by renting households
  - Percent high rent: Percent of renting households paying \$2,000 or more per month
  - Percent owning: Percent of households who own their home
  - Median house value: Median value of house among owning households
  - Percent high house value: Percent of owning households with house values of \$500,000 or more

#### Housing count

```{r hmap}
# total housing
breaks <- c(0,1,500,1000,2000,2500)
legend_labels <- c("None", "500 or less", "501 to 1000", "1001 to 2000", "2001 or more")
sel_range <- c(min(breaks), max(breaks))

pal <- colorBin("Purples", sel_range, bins = breaks, right = TRUE, reverse = FALSE, na.color = "#808080", pretty = FALSE )

hh_total <- leaflet(bg_full) %>%
  addProviderTiles('CartoDB.Positron') %>%
  addPolygons(data = bg_full,
              weight = 0.5,
              color = "#FFFFFF",
              smoothFactor = 0.2,
              fillColor = ~pal(hh_totalE),
              fillOpacity = 0.7,
              popup = paste0("Value: ", round(bg_full$hh_totalE, 1))) %>%
  addLegend(position = 'bottomright', 
            pal = pal,
            values = sel_range,
            labFormat = function(type, breaks) {
              return(legend_labels)
            },
            title = "Total Housing Units",
            opacity = 0.7)

hh_total
```

#### Percent vacant

```{r hmap2}
# Vacant
breaks <- c(0,10,20,40,60,80,100)
legend_labels <- c("10% or less", "10% to 20%", "20% to 40%", "40% to 60%", "60% to 80%", "80% to 100%")
sel_range <- c(min(breaks), max(breaks))

pal <- colorBin("Purples", sel_range, bins = breaks, right = TRUE, reverse = FALSE, na.color = "#808080", pretty = FALSE )

hh_vacant <- leaflet(bg_full) %>%
  addProviderTiles('CartoDB.Positron') %>%
  addPolygons(data = bg_full,
              weight = 0.5,
              color = "#FFFFFF",
              smoothFactor = 0.2,
              fillColor = ~pal(hh_per_vacant),
              fillOpacity = 0.7,
              popup = paste0("Value: ", round(bg_full$hh_per_vacant, 1))) %>%
  addLegend(position = 'bottomright', 
            pal = pal,
            values = sel_range,
            labFormat = function(type, breaks) {
              return(legend_labels)
            },
            title = "Percent Vacant Housing",
            opacity = 0.7)

hh_vacant
```

#### Median Income

```{r hmap3}
# median income
breaks <- c(0,30000,50000,75000,100000,125000)
legend_labels <- c("$30K or less", "$30K to $50K", "$50K to $75K", "$75K to $100K", "$100K or more")
sel_range <- c(min(breaks), max(breaks))

pal <- colorBin("Purples", sel_range, bins = breaks, right = TRUE, reverse = FALSE, na.color = "#808080", pretty = FALSE )

hh_medinc <- leaflet(bg_full) %>%
  addProviderTiles('CartoDB.Positron') %>%
  addPolygons(data = bg_full,
              weight = 0.5,
              color = "#FFFFFF",
              smoothFactor = 0.2,
              fillColor = ~pal(hh_medincE),
              fillOpacity = 0.7,
              popup = paste0("Value: ", round(bg_full$hh_medincE, 1))) %>%
  addLegend(position = 'bottomright', 
            pal = pal,
            values = sel_range,
            labFormat = function(type, breaks) {
              return(legend_labels)
            },
            title = "Median Household Income",
            opacity = 0.7)

hh_medinc
```

#### Percent high income

```{r hmap4}
# high income
breaks <- c(0,10,20,40,60,80,100)
legend_labels <- c("10% or less", "10% to 20%", "20% to 40%", "40% to 60%", "60% to 80%", "80% to 100%")
sel_range <- c(min(breaks), max(breaks))

pal <- colorBin("Purples", sel_range, bins = breaks, right = TRUE, reverse = FALSE, na.color = "#808080", pretty = FALSE )

hh_hiinc <- leaflet(bg_full) %>%
  addProviderTiles('CartoDB.Positron') %>%
  addPolygons(data = bg_full,
              weight = 0.5,
              color = "#FFFFFF",
              smoothFactor = 0.2,
              fillColor = ~pal(hh_per_hhinc_200K),
              fillOpacity = 0.7,
              popup = paste0("Value: ", round(bg_full$hh_per_hhinc_200K, 1))) %>%
  addLegend(position = 'bottomright', 
            pal = pal,
            values = sel_range,
            labFormat = function(type, breaks) {
              return(legend_labels)
            },
            title = "Percent with Income $200K or more",
            opacity = 0.7)

hh_hiinc
```

#### Percent renting

```{r hmap5}
# Percent rent
breaks <- c(0,10,20,40,60,80,100)
legend_labels <- c("10% or less", "10% to 20%", "20% to 40%", "40% to 60%", "60% to 80%", "80% to 100%")
sel_range <- c(min(breaks), max(breaks))

pal <- colorBin("Purples", sel_range, bins = breaks, right = TRUE, reverse = FALSE, na.color = "#808080", pretty = FALSE )

hh_rent <- leaflet(bg_full) %>%
  addProviderTiles('CartoDB.Positron') %>%
  addPolygons(data = bg_full,
              weight = 0.5,
              color = "#FFFFFF",
              smoothFactor = 0.2,
              fillColor = ~pal(hh_per_rent),
              fillOpacity = 0.7,
              popup = paste0("Value: ", round(bg_full$hh_per_rent, 1))) %>%
  addLegend(position = 'bottomright', 
            pal = pal,
            values = sel_range,
            labFormat = function(type, breaks) {
              return(legend_labels)
            },
            title = "Percent Renting",
            opacity = 0.7)

hh_rent
```

#### Median rent

```{r hmap6}
# median rent
breaks <- c(0,600,800,1000,1200,2000)
legend_labels <- c("$600 or less", "$600 to $800", "$800 to $1,000", "$1,000 to $1,200", "$1,200 or more")
sel_range <- c(min(breaks), max(breaks))

pal <- colorBin("Purples", sel_range, bins = breaks, right = TRUE, reverse = FALSE, na.color = "#808080", pretty = FALSE )

hh_medrent <- leaflet(bg_full) %>%
  addProviderTiles('CartoDB.Positron') %>%
  addPolygons(data = bg_full,
              weight = 0.5,
              color = "#FFFFFF",
              smoothFactor = 0.2,
              fillColor = ~pal(hh_rented_medrentE),
              fillOpacity = 0.7,
              popup = paste0("Value: ", round(bg_full$hh_rented_medrentE, 1))) %>%
  addLegend(position = 'bottomright', 
            pal = pal,
            values = sel_range,
            labFormat = function(type, breaks) {
              return(legend_labels)
            },
            title = "Median Monthly Rent",
            opacity = 0.7)

hh_medrent
```

#### Percent high rent

```{r hmap7}
# Percent high rent
breaks <- c(0,10,20,40,60,80,100)
legend_labels <- c("10% or less", "10% to 20%", "20% to 40%", "40% to 60%", "60% to 80%", "80% to 100%")
sel_range <- c(min(breaks), max(breaks))

pal <- colorBin("Purples", sel_range, bins = breaks, right = TRUE, reverse = FALSE, na.color = "#808080", pretty = FALSE )

hh_rent2k <- leaflet(bg_full) %>%
  addProviderTiles('CartoDB.Positron') %>%
  addPolygons(data = bg_full,
              weight = 0.5,
              color = "#FFFFFF",
              smoothFactor = 0.2,
              fillColor = ~pal(hh_per_rented_2k),
              fillOpacity = 0.7,
              popup = paste0("Value: ", round(bg_full$hh_per_rented_2k, 1))) %>%
  addLegend(position = 'bottomright', 
            pal = pal,
            values = sel_range,
            labFormat = function(type, breaks) {
              return(legend_labels)
            },
            title = "Percent Paying $2K or more",
            opacity = 0.7)

hh_rent2k
```

#### Percent owning

```{r hmap8}
# Percent own
breaks <- c(0,10,20,40,60,80,100)
legend_labels <- c("10% or less", "10% to 20%", "20% to 40%", "40% to 60%", "60% to 80%", "80% to 100%")
sel_range <- c(min(breaks), max(breaks))

pal <- colorBin("Purples", sel_range, bins = breaks, right = TRUE, reverse = FALSE, na.color = "#808080", pretty = FALSE )

bg_full <- bg_full %>% 
  mutate(hh_per_own = 100-hh_per_rent)

hh_own <- leaflet(bg_full) %>%
  addProviderTiles('CartoDB.Positron') %>%
  addPolygons(data = bg_full,
              weight = 0.5,
              color = "#FFFFFF",
              smoothFactor = 0.2,
              fillColor = ~pal(hh_per_own),
              fillOpacity = 0.7,
              popup = paste0("Value: ", round(bg_full$hh_per_own, 1))) %>%
  addLegend(position = 'bottomright', 
            pal = pal,
            values = sel_range,
            labFormat = function(type, breaks) {
              return(legend_labels)
            },
            title = "Percent Owning",
            opacity = 0.7)

hh_own
```

#### Median house value

```{r hmap9}
# median house vale
breaks <- c(0,100000,150000,250000,350000,600000)
legend_labels <- c("$100K or less", "$100K to $150K", "$150K to $250K", "$250K to $350K", "$350K or more")
sel_range <- c(min(breaks), max(breaks))

pal <- colorBin("Purples", sel_range, bins = breaks, right = TRUE, reverse = FALSE, na.color = "#808080", pretty = FALSE )

hh_medvalue <- leaflet(bg_full) %>%
  addProviderTiles('CartoDB.Positron') %>%
  addPolygons(data = bg_full,
              weight = 0.5,
              color = "#FFFFFF",
              smoothFactor = 0.2,
              fillColor = ~pal(hh_owned_medvalueE),
              fillOpacity = 0.7,
              popup = paste0("Value: ", round(bg_full$hh_owned_medvalueE, 1))) %>%
  addLegend(position = 'bottomright', 
            pal = pal,
            values = sel_range,
            labFormat = function(type, breaks) {
              return(legend_labels)
            },
            title = "Median Home Value",
            opacity = 0.7)

hh_medvalue
```

#### Percent high house value

```{r hmap10}
# Percent high house value
breaks <- c(0,10,20,40,60,80,100)
legend_labels <- c("10% or less", "10% to 20%", "20% to 40%", "40% to 60%", "60% to 80%", "80% to 100%")
sel_range <- c(min(breaks), max(breaks))

pal <- colorBin("Purples", sel_range, bins = breaks, right = TRUE, reverse = FALSE, na.color = "#808080", pretty = FALSE )

hh_ownn500k <- leaflet(bg_full) %>%
  addProviderTiles('CartoDB.Positron') %>%
  addPolygons(data = bg_full,
              weight = 0.5,
              color = "#FFFFFF",
              smoothFactor = 0.2,
              fillColor = ~pal(hh_per_owned_500k),
              fillOpacity = 0.7,
              popup = paste0("Value: ", round(bg_full$hh_per_owned_500k, 1))) %>%
  addLegend(position = 'bottomright', 
            pal = pal,
            values = sel_range,
            labFormat = function(type, breaks) {
              return(legend_labels)
            },
            title = "Percent Homes Valued $500K or more",
            opacity = 0.7)

hh_ownn500k
```
