---
title: "ESVA Population Summaries"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r prep}
library(sf)
library(jsonlite)
library(tidyverse)
library(patchwork)
library(leaflet)

### Blocks
blocks <- st_read("cb/esva_2020block_clipped.geojson",
                  quiet = TRUE)
blocks <- st_transform(blocks, 4326)

### Population
pop <- read_csv("block_pop_data.csv") %>% 
  mutate(GEOID = as.character(GEOID))

### Storm surge
ea_export <- fromJSON("ADCIRC_stormsurge/esva_Block20_IB2003_PeakWL_ea_exportV2.json") # not reading
json <- fromJSON("ADCIRC_stormsurge/esva_Block20_IB2003_PeakWLV2.json")
storm <- read_csv("ADCIRC_stormsurge/esva_Block20_IB2003_PeakWLV2.csv") %>% 
  mutate(GEOID20 = as.character(GEOID20))

## Combine data
df <- left_join(blocks, pop, by=c("GEOID20"="GEOID")) %>%
  left_join(storm, by = c("GEOID20"="GEOID20")) %>% 
  st_drop_geometry()

df_map <- left_join(blocks, storm, by = c("GEOID20"="GEOID20"))

### Total population baselines
pop_total <- pop %>% 
  summarize(across(c(total:jobs_wac_hi), sum)) %>% 
  mutate(across(c(hisp:pop_under18), ~ (.x/total)*100, .names = "per_{.col}"),
         per_occupied_housing = (occupied_housing/total_housing)*100,
         across(c(jobs_rac_low:jobs_rac_hi), ~ (.x/jobs_rac)*100, .names = "per_{.col}"),
         across(c(jobs_wac_low:jobs_wac_hi), ~ (.x/jobs_wac)*100, .names = "per_{.col}"),
         bin = "ESVA Total", event = "none")

```

## Map

```{r map}
# From Beth's leaflet
region_bbox <- ea_export$regionBoundingBox
bbox <- c(region_bbox$lonMin, region_bbox$latMin, region_bbox$lonMax, region_bbox$latMax)

legend_breaks <- as.vector(ea_export$dataColumns$bins[[1]])
legend_labels <- as.list(ea_export$dataColumns$labels[[1]])

descriptionTitle <- ea_export$descriptionTitle

# var1_name <- ea_export$dataColumns$name[1]
var1_name = "MxWL_avg_ft"
var1_name_index <- match(var1_name, json$schema$fields$name)
var1_title <- json$schema$fields$title[var1_name_index]

var2_name <- ea_export$dataColumns$name[2]
var2_name_index <- match(var2_name, json$schema$fields$name)
var2_title <- json$schema$fields$title[var2_name_index]

m <- df_map

var1 <- m[[var1_name]]
var2 <- m[[var2_name]]

breaks <- legend_breaks
sel_range <- c(min(breaks), max(breaks))

pal <- colorBin(c("#FEF0D9", "#FDD49E", "#FDBB84", "#FC8D59", "#E34A33", "#B30000"), sel_range, bins = breaks, right = TRUE, reverse = FALSE, na.color = "#808080", pretty = FALSE )

map <- leaflet(m) %>%
  addProviderTiles('CartoDB.Positron') %>%
  fitBounds(bbox[1], bbox[2], bbox[3], bbox[4]) %>% 
  clearShapes() %>%
  addPolygons(data = m,
              weight = 0.5,
              color = "#FFFFFF",
              smoothFactor = 0.2,
              fillColor = ~pal(var1),
              fillOpacity = 0.7,
              group = var1_title) %>%
  addPolygons(data = m,
              weight = 0.5,
              color = "#FFFFFF",
              smoothFactor = 0.2, 
              fillColor = ~pal(var2),
              fillOpacity = 0.7,
              group = var2_title) %>%
  addLegend(position = 'bottomright', 
            pal = pal,
            values = sel_range,
            labFormat = function(type, breaks) {
              return(legend_labels)
            },
            title = ~gsub("\n", "<br>",
                          stringr::str_wrap(descriptionTitle,
                                            width = 20,
                                            whitespace_only = FALSE)),
            opacity = 0.7) %>% 
  addLayersControl(
    baseGroups = c(var1_title, var2_title),
    options = layersControlOptions(collapsed = FALSE)
  ) 

map
```

```{r popfig}
df <- df %>% 
  mutate(bin = cut(MxWL_max_ft, 
                     breaks = legend_breaks,
                     labels = legend_labels,
                     include.lowest = TRUE),
         event = "stormsurge") 

## Impact estimates
df_impact <- df %>% 
  group_by(bin, event) %>% 
  summarize(across(c(total:jobs_wac_hi), sum)) %>% 
  mutate(across(c(hisp:pop_under18), ~ (.x/total)*100, .names = "per_{.col}"),
         per_occupied_housing = (occupied_housing/total_housing)*100,
         across(c(jobs_rac_low:jobs_rac_hi), ~ (.x/jobs_rac)*100, .names = "per_{.col}"),
         across(c(jobs_wac_low:jobs_wac_hi), ~ (.x/jobs_wac)*100, .names = "per_{.col}")) %>% 
  ungroup() %>% 
  mutate(per_total = (total/sum(total))*100,
         per_housing = (total_housing/sum(total_housing))*100) 
  
df_impact_total <- bind_rows(df_impact, pop_total) %>% 
  mutate(bin = factor(bin),
         bin = fct_relevel(bin, "ESVA Total", after = 0))
```

## Population Impacts of Peak Surge

### Distribution of overall population and housing

```{r viz, fig.height=3}
pal <- c("#FEF0D9", "#FDD49E", "#FDBB84", "#FC8D59", "#E34A33", "#B30000")

popvars <- select(df_impact_total, starts_with("per")) %>% names()
poplabels <- c("Hispanic", "White", "Black", "American Indian/Alaskan Native",
               "Asian", "Native Hawaiian/Pacific Islander", "Remaining Racial Categories",
               "Multiracial", "Over 18", "Under 18", "Occupied Housing",
               "By Residence", "Mid-wage (Residence)",
               "High-wage (Residence)", "By Workplace",
               "Mid-wage (Workplace)", "High-wage (Workplace)",
               "Total Population", "Total Housing Stock")

storm_plots_bar_tot <- map2(popvars[18:19], poplabels[18:19], 
                    ~ df_impact_total %>% 
                      filter(event == "stormsurge") %>% 
                      ggplot(aes(x = bin, y = .data[[.x]])) +
                      geom_col(aes(fill = bin)) +
                      geom_text(aes(y = .data[[.x]], 
                                    label = round(.data[[.x]], 1)),
                                vjust = 1, color = "black") +
                      scale_fill_manual(values = pal, na.value="grey", guide = "none") +
                      labs(title = paste0("Percent of: ", .y),
                           subtitle = "By Impact",
                           x = "", y = "") +
                      theme_classic())

wrap_plots(storm_plots_bar_tot, ncol = 2)

```

```{r viz2, fig.height=3}
storm_plots_bar <- map2(popvars[c(2,3,1,8,5,4,7,6,9,10,12,15,13,14,16,17)], poplabels[c(2,3,1,8,5,4,7,6,9,10,12,15,13,14,16,17)], 
                    ~ df_impact_total %>% 
                     filter(event == "stormsurge") %>% 
                     ggplot(aes(x = bin, y = .data[[.x]])) +
                     geom_col(aes(fill = bin)) +
                     geom_text(aes(y = .data[[.x]], 
                                   label = round(.data[[.x]], 1)),
                               vjust = 1, color = "black") +
  geom_hline(aes(yintercept = df_impact_total %>% 
                   filter(bin == "ESVA Total") %>% 
                   pull(.x)), color = "brown4", linetype = 2) +
    expand_limits(x= c(1, 8)) +
    annotate("text", x = 7, label = "ESVA\nTotal", color = "brown4",
             y = df_impact_total %>% 
               filter(event == "none") %>% 
               pull(.x)) +   
    scale_fill_manual(values = pal, na.value="grey", guide = "none") +
    labs(title = paste0("Percent of Residents: ", .y),
       subtitle = "By Impact",
       x = "", y = "") +
  theme_classic())
```

### Distribution of population by Racial Identity

#### Racial identity categories with > 500 estimated residents 

```{r, fig.height=7}
wrap_plots(storm_plots_bar[1:4], ncol = 2)
```

#### Racial identity categories with < 200 estimated residents

```{r, fig.height=7}
wrap_plots(storm_plots_bar[5:8], ncol = 2)
```

### Distribution of population by Age

```{r, fig.height=3}
wrap_plots(storm_plots_bar[9:10], ncol = 2)
```

### Distribution of population by Low-Wage Workers

Earning $1,250/month or less

```{r, fig.height=3}
wrap_plots(storm_plots_bar[11:12], ncol = 2)
```

#### Distribution of Other Workers by Residence

Mid-wage: $1,251-$3,333/month; High-wage: More than $3,333/month

```{r, fig.height=3}
wrap_plots(storm_plots_bar[13:14], ncol = 2)
```

#### Distribution of Other Workers by Workplace

Mid-wage: $1,251-$3,333/month; High-wage: More than $3,333/month

```{r, fig.height=3}
wrap_plots(storm_plots_bar[15:16], ncol = 2)
```
