---
title: "ESVA Population Impact - Storm Surge Risk"
output: html_document
date: 2025-10-31
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r prep}
library(sf)
library(tidyverse)
library(patchwork)
library(leaflet)

pal <- RColorBrewer::brewer.pal(5, "Purples")

# Data
# Created in poptest4.R
load("ssr_pop_example.Rdata")

# block shapefile
blocks <- st_read("cb/esva_2020block_clipped.geojson",
                  quiet = TRUE)
blocks <- st_transform(blocks, 4326)

## Combine data for maps
ssr_df_map <- left_join(blocks, ssr_df, by = c("GEOID20"="GEOID20"))
```

## Map of Storm Surge Risk

```{r map, fig.width=10}
m <- ssr_df_map

var1 <- m[[var1_name]]

breaks <- legend_breaks
sel_range <- c(min(breaks), max(breaks))

pal <- colorBin("Purples", sel_range, bins = breaks, right = TRUE, reverse = FALSE, na.color = "#808080", pretty = FALSE )

map <- leaflet(m) %>%
  addProviderTiles('CartoDB.Positron') %>%
  fitBounds(bbox[1], bbox[2], bbox[3], bbox[4]) %>% 
  clearShapes() %>%
  addPolygons(data = m,
              weight = 0.5,
              color = "#FFFFFF",
              smoothFactor = 0.2,
              fillColor = ~pal(var1),
              fillOpacity = 0.7,
              group = var1_title) %>%
  addLegend(position = 'bottomright', 
            pal = pal,
            values = sel_range,
            labFormat = function(type, breaks) {
              return(legend_labels)
            },
            title = ~gsub("\n", "<br>",
                          stringr::str_wrap(descriptionTitle,
                                            width = 20,
                                            whitespace_only = FALSE)),
            opacity = 0.7)

map
```

## Population Impacts of Storm Surge Risk

### Distribution of blocks, housing units, population

```{r viz1, fig.height=3, fig.width=10}
wrap_plots(bar_plots_tot, ncol = 3)
```


### Distribution by Racial & Ethnic Identity

Racial identity categories with > 1,000 estimated residents 

#### Composition

Percent of residents in target population within each outcome range

```{r viz2, fig.height=7, fig.width=10}
wrap_plots(bar_plots[c(2,3,1,8)], ncol = 2)
```

#### Risk Ratio

Ratio of risk for targeted group relative to all other residents

```{r viz2b, fig.height=7, fig.width=10}
wrap_plots(bar_plots_rr[c(2,3,1,8)], ncol = 2, guides = "collect")
```


### Distribution by Age

#### Composition

Percent of residents in target population within each outcome range

```{r viz3, fig.height=3, fig.width=10}
wrap_plots(bar_plots[c(10,9)], ncol = 2)
```

#### Risk Ratio

Ratio of risk for targeted group relative to all other residents

```{r viz3b, fig.height=3, fig.width=10}
wrap_plots(bar_plots_rr[c(10,9)], ncol = 2, guides = "collect")
```

### Distribution among Low-Wage Workers

By Place of Residence, By Place of Work

* Low-Wage: Earning $15,000/year or less

#### Composition 

Percent of residents in target population within each outcome range

```{r viz4, fig.height=3, fig.width=10}
wrap_plots(bar_plots[c(12,15)], ncol = 2)
```

#### Risk Ratio

Ratio of risk for targeted group relative to all other residents

```{r viz4b, fig.height=3, fig.width=10}
wrap_plots(bar_plots_rr[c(11,14)], ncol = 2, guides = "collect")
```
