---
title: "ESVA Population Impact Example"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r prep}
library(sf)
library(tidyverse)
library(patchwork)
library(leaflet)

# Data
# Created in poptest2.R
load("gw_pop_example.Rdata")

# block shapefile
blocks <- st_read("cb/esva_2020block_clipped.geojson",
                  quiet = TRUE)
blocks <- st_transform(blocks, 4326)

## Combine data for maps
gw_df_map <- left_join(blocks, gw_df, by = c("GEOID20"="GEOID20"))
```

## Map of Depth to Groundwater (ft) - 2030 (Average)

```{r map, fig.width=10}
m <- gw_df_map

var1 <- m[[var1_name]]

breaks <- legend_breaks
sel_range <- c(min(breaks), max(breaks))

pal <- colorBin(c("#FEF0D9", "#FDD49E", "#FDBB84", "#FC8D59", "#E34A33", "#B30000"), sel_range, bins = breaks, right = TRUE, reverse = FALSE, na.color = "#808080", pretty = FALSE )

map <- leaflet(m) %>%
  addProviderTiles('CartoDB.Positron') %>%
  fitBounds(bbox[1], bbox[2], bbox[3], bbox[4]) %>% 
  clearShapes() %>%
  addPolygons(data = m,
              weight = 0.5,
              color = "#FFFFFF",
              smoothFactor = 0.2,
              fillColor = ~pal(var1),
              fillOpacity = 0.7,
              group = var1_title) %>%
  addLegend(position = 'bottomright', 
            pal = pal,
            values = sel_range,
            labFormat = function(type, breaks) {
              return(legend_labels)
            },
            title = ~gsub("\n", "<br>",
                          stringr::str_wrap(descriptionTitle,
                                            width = 20,
                                            whitespace_only = FALSE)),
            opacity = 0.7)

map
```

## Population Impacts of Depth to Groundwater (ft) - 2030 (Average)

```{r makeviz}
pal <- c("#FEF0D9", "#FDD49E", "#FDBB84", "#FC8D59", "#E34A33", "#B30000")

bar_plots <- map2(popvars[1:17], poplabels[1:17], 
                        ~ gw_df_test %>% 
                          filter(event != "none") %>% 
                          ggplot(aes(x = bin, y = .data[[.x]])) +
                          geom_col(aes(fill = bin)) +
                          geom_text(aes(y = .data[[.x]], 
                                        label = round(.data[[.x]], 1)),
                                    vjust = 1, color = "black") +
                          geom_hline(aes(yintercept = gw_df_test %>% 
                                           filter(bin == "ESVA Total") %>% 
                                           pull(.x)), color = "brown4", linetype = 2) +
                          expand_limits(x= c(1, 8)) +
                          annotate("text", x = 7, label = "ESVA\nTotal", color = "brown4",
                                   y = gw_df_test %>% 
                                     filter(event == "none") %>% 
                                     pull(.x)) +   
                          scale_fill_manual(values = pal, na.value="grey", guide = "none") +
                          labs(title = paste0("Percent of Residents: ", .y),
                               subtitle = "By Impact",
                               x = "", y = "") +
                          theme_classic())

bar_plots_tot <- map2(popvars[18:19], poplabels[18:19], 
                            ~ gw_df_test %>% 
                              filter(event != "none") %>% 
                              ggplot(aes(x = bin, y = .data[[.x]])) +
                              geom_col(aes(fill = bin)) +
                              geom_text(aes(y = .data[[.x]], 
                                            label = round(.data[[.x]], 1)),
                                        vjust = 1, color = "black") +
                              scale_fill_manual(values = pal, na.value="grey", guide = "none") +
                              labs(title = paste0("Percent of: ", .y),
                                   subtitle = "By Impact",
                                   x = "", y = "") +
                              theme_classic())
```


### Distribution of overall population and housing

```{r viz1, fig.height=3, fig.width=10}
wrap_plots(bar_plots_tot, ncol = 2)
```

### Distribution by Racial & Ethnic Identity

#### Racial identity categories with > 1,000 estimated residents 

```{r viz2, fig.height=7, fig.width=10}
wrap_plots(bar_plots[c(2,3,1,8)], ncol = 2)
```

### Distribution by Age

```{r, fig.height=3, fig.width=10}
wrap_plots(bar_plots[c(10,9)], ncol = 2)
```

### Distribution among Low-Wage Workers

By Place of Residence, By Place of Work

* Low-Wage: Earning $1,250/month or less

```{r, fig.height=3, fig.width=10}
wrap_plots(bar_plots[c(12,15)], ncol = 2)
```

