---
title: "ESVA: Who are you worried about?"
date: "2024-02-29"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r libs}
library(tidyverse)
library(scales)
library(readxl)
library(sf)
library(ggiraph) # https://www.ardata.fr/ggiraph-book/
library(patchwork)
library(ggspatial) # appears to require installation of "rosm" and "prettymapr" as well
```

```{r data}
# flood hazard data
flood_blkgrp <- readRDS("../flood_composite_blkgrp.RDS")

# make 10 percent columns
flood_blkgrp <- flood_blkgrp %>% 
  mutate(across(starts_with("sum"), ~.x/cells, .names = "per_{.col}"))

# pop data
pop_blkgrp <- read_csv("../population_blkgrp.csv")
pop_blkgrp <- pop_blkgrp %>% 
  separate_wider_delim(NAME, delim = "; ", 
                       names = c("block_grp", "tract", "county", "state")) %>% 
  mutate(GEOID = as.character(GEOID))

# block group names
blkgrp_names <- read_excel("../tract_names.xlsx", sheet = "blkgrp2020")
blkgrp_names <- blkgrp_names %>% 
  mutate(localityfips = str_pad(localityfips, width = 3, side = "left", pad = "0"),
         tract = str_pad(tract, width = 6, side = "left", pad = "0"),
         GEOID = paste0("51",localityfips,tract,blkgrp),
         names = str_remove(names, "'"))

# join
bg_df <- flood_blkgrp %>% 
  left_join(pop_blkgrp, by = "GEOID") %>% 
  left_join(blkgrp_names, by = "GEOID") 
```

## NOAA Coastal Flood Hazard Components

### Percent of area subject to high tide flooding

```{r flood1}
# change hover info
bg_df <- bg_df %>% 
  mutate(toolinfo = paste0(names, "\n", 
                           "Percent: ",
                           round(per_sum_HighTideFlooding*100, 0), "%"))

# make map
gg_map <- ggplot(bg_df) +
  # annotation_map_tile(type = "cartolight") +
  geom_sf_interactive(aes(fill = per_sum_HighTideFlooding, 
                          tooltip = toolinfo, 
                          data_id = GEOID), 
                      color = "white") +
  scale_fill_gradient(low = "grey90", high = "darkblue", 
                      name = "Percent\nof Area Subject\nto High Tide\nFlooding", labels = scales::percent_format(scale = 100)) +
  theme_void() +
  theme(legend.title = element_text(size = 6),
        legend.text = element_text(size = 5),
        legend.position = c(.8, .4))

# make points
gg_point <- ggplot(bg_df, aes(x = per_sum_HighTideFlooding, 
                              y = fct_rev(county), 
                              tooltip = toolinfo, 
                              data_id = GEOID)) +
  geom_point_interactive(size = 7, alpha = 0.2, color = "#98817b") +
  scale_x_continuous(limits = c(0, 1), 
                     labels = scales::percent_format(scale = 100), name = "Percent of Area") +
  scale_y_discrete(name = "", 
                   labels = function(groups) str_wrap(groups, width = 10)) +
  theme_minimal() +
  theme(legend.title=element_blank(),
        legend.text = element_text(size = 12),
        axis.text.x = element_text(size = 11),
        strip.background = element_rect(color = "black", fill = "#f1f1f1"),
        strip.text.x = element_text(size = 12),
        strip.placement = "outside")

girafe(ggobj = gg_point + gg_map + plot_layout(widths = c(0.45, 0.65)), 
       width_svg = 12, height_svg = 6) %>%
  girafe_options(opts_hover(css = "fill:darkorange;stroke:darkblue;stroke-width:2px;"))         
```

### Percent of area subject to storm surge (catgory 2) flooding

```{r flood2}
# change hover info
bg_df <- bg_df %>% 
  mutate(toolinfo = paste0(names, "\n", 
                           "Percent: ",
                           round(per_sum_ss.cat2*100, 0), "%"))

# make map
gg_map <- ggplot(bg_df) +
  # annotation_map_tile(type = "cartolight") +
  geom_sf_interactive(aes(fill = per_sum_ss.cat2,
                          tooltip = toolinfo, data_id = GEOID), 
                      color = "white") +
  scale_fill_gradient(low = "grey90", high = "darkblue", 
                      name = "Percent\nof Area Subject\nto Storm Surge\nFlooding (Cat 2)", labels = scales::percent_format(scale = 100)) +
  theme_void() +
  theme(legend.title = element_text(size = 6),
        legend.text = element_text(size = 5),
        legend.position = c(.8, .4))

# make points
gg_point <- ggplot(bg_df, aes(x = per_sum_ss.cat2, 
                                   y = fct_rev(county),
                                   tooltip = toolinfo, 
                                   data_id = GEOID)) +
  geom_point_interactive(size = 7, alpha = 0.2, color = "#98817b") +
  scale_x_continuous(limits = c(0, 1),
                     labels = scales::percent_format(scale = 100),
                     name = "Percent of Area") +
  scale_y_discrete(name = "",
                   labels = function(groups) str_wrap(groups, width = 10)) +
  theme_minimal() +
  theme(legend.title=element_blank(),
        legend.text = element_text(size = 12),
        axis.text.x = element_text(size = 11),
        strip.background = element_rect(color = "black", fill = "#f1f1f1"),
        strip.text.x = element_text(size = 12),
        strip.placement = "outside")

girafe(ggobj = gg_point + gg_map + plot_layout(widths = c(0.45, 0.65)), 
       width_svg = 12, height_svg = 6) %>%
  girafe_options(opts_hover(css = "fill:darkorange;stroke:darkblue;stroke-width:2px;"))         
```

### Percent of area subject to flooding from sea level rise (2 feet)

```{r flood3}
# change hover info
bg_df <- bg_df %>% 
  mutate(toolinfo = paste0(names, "\n", 
                           "Percent: ",
                           round(per_sum_slr2.ft*100, 0), "%"))

# make map
gg_map <- ggplot(bg_df) +
  # annotation_map_tile(type = "cartolight") +
  geom_sf_interactive(aes(fill = per_sum_slr2.ft,
                          tooltip = toolinfo, data_id = GEOID), 
                      color = "white") +
  scale_fill_gradient(low = "grey90", high = "darkblue", 
                      name = "Percent of\nArea Subject to\nSea Level Rise (2ft)", labels = scales::percent_format(scale = 100)) +
  theme_void() +
  theme(legend.title = element_text(size = 6),
        legend.text = element_text(size = 5),
        legend.position = c(.8, .4))

# make points
gg_point <- ggplot(bg_df, aes(x = per_sum_slr2.ft, 
                                   y = fct_rev(county),
                                   tooltip = toolinfo, 
                                   data_id = GEOID)) +
  geom_point_interactive(size = 7, alpha = 0.2, color = "#98817b") +
  scale_x_continuous(limits = c(0, 1),
                     labels = scales::percent_format(scale = 100),
                     name = "Percent of Area") +
  scale_y_discrete(name = "",
                   labels = function(groups) str_wrap(groups, width = 10)) +
  theme_minimal() +
  theme(legend.title=element_blank(),
        legend.text = element_text(size = 12),
        axis.text.x = element_text(size = 11),
        strip.background = element_rect(color = "black", fill = "#f1f1f1"),
        strip.text.x = element_text(size = 12),
        strip.placement = "outside")

girafe(ggobj = gg_point + gg_map + plot_layout(widths = c(0.45, 0.65)), 
       width_svg = 12, height_svg = 6) %>%
  girafe_options(opts_hover(css = "fill:darkorange;stroke:darkblue;stroke-width:2px;"))         
```

### Percent of area in 2% annual chance flood zone

```{r flood4}
# change hover info
bg_df <- bg_df %>% 
  mutate(toolinfo = paste0(names, "\n", 
                           "Percent: ",
                           round(per_sum_azone.2percent*100, 0), "%"))

# make map
gg_map <- ggplot(bg_df) +
  # annotation_map_tile(type = "cartolight") +
  geom_sf_interactive(aes(fill = per_sum_azone.2percent,
                          tooltip = toolinfo, data_id = GEOID), 
                      color = "white") +
  scale_fill_gradient(low = "grey90", high = "darkblue", 
                      name = "Percent of\nArea in the 2%\nAnnual Flood Zone", labels = scales::percent_format(scale = 100)) +
  theme_void() +
  theme(legend.title = element_text(size = 6),
        legend.text = element_text(size = 5),
        legend.position = c(.8, .4))

# make points
gg_point <- ggplot(bg_df, aes(x = per_sum_azone.2percent, 
                                   y = fct_rev(county),
                                   tooltip = toolinfo, 
                                   data_id = GEOID)) +
  geom_point_interactive(size = 7, alpha = 0.2, color = "#98817b") +
  scale_x_continuous(limits = c(0, 1),
                     labels = scales::percent_format(scale = 100),
                     name = "Percent of Area") +
  scale_y_discrete(name = "",
                   labels = function(groups) str_wrap(groups, width = 10)) +
  theme_minimal() +
  theme(legend.title=element_blank(),
        legend.text = element_text(size = 12),
        axis.text.x = element_text(size = 11),
        strip.background = element_rect(color = "black", fill = "#f1f1f1"),
        strip.text.x = element_text(size = 12),
        strip.placement = "outside")

girafe(ggobj = gg_point + gg_map + plot_layout(widths = c(0.45, 0.65)), 
       width_svg = 12, height_svg = 6) %>%
  girafe_options(opts_hover(css = "fill:darkorange;stroke:darkblue;stroke-width:2px;"))         
```

### Average flood exposure composite across four flood measures

```{r flood5}
# change hover info
bg_df <- bg_df %>% 
  mutate(toolinfo = paste0(names, "\n", 
                           "Average: ",
                           round(mean_HazardNumber, 1)))

# make map
gg_map <- ggplot(bg_df) +
  # annotation_map_tile(type = "cartolight") +
  geom_sf_interactive(aes(fill = mean_HazardNumber,
                          tooltip = toolinfo, data_id = GEOID), 
                      color = "white") +
  scale_fill_gradient(low = "grey90", high = "darkblue", 
                      name = "Average\nFlood\nHazard\nCompsite") +
  theme_void() +
  theme(legend.title = element_text(size = 6),
        legend.text = element_text(size = 5),
        legend.position = c(.8, .4))

# make points
gg_point <- ggplot(bg_df, aes(x = mean_HazardNumber, 
                                   y = fct_rev(county),
                                   tooltip = toolinfo, 
                                   data_id = GEOID)) +
  geom_point_interactive(size = 7, alpha = 0.2, color = "#98817b") +
  scale_x_continuous(limits = c(0, 1),
                     labels = scales::percent_format(scale = 100),
                     name = "Average across Area") +
  scale_y_discrete(name = "",
                   labels = function(groups) str_wrap(groups, width = 10)) +
  theme_minimal() +
  theme(legend.title=element_blank(),
        legend.text = element_text(size = 12),
        axis.text.x = element_text(size = 11),
        strip.background = element_rect(color = "black", fill = "#f1f1f1"),
        strip.text.x = element_text(size = 12),
        strip.placement = "outside")

girafe(ggobj = gg_point + gg_map + plot_layout(widths = c(0.45, 0.65)), 
       width_svg = 12, height_svg = 6) %>%
  girafe_options(opts_hover(css = "fill:darkorange;stroke:darkblue;stroke-width:2px;"))         
```

## Population Characteristics

### Workers in jobs paying less than $15,000/year, by place of work
```{r pop1}
# change hover info
bg_df <- bg_df %>% 
  mutate(toolinfo = paste0(names, "\n", 
                           "Percent: ",
                           round(percent_lowage_w, 1), "%",
                           "\n", "Number: ", jobs_lowage_w))

# make map
gg_map <- ggplot(bg_df) +
  # annotation_map_tile(type = "cartolight") +
  geom_sf_interactive(aes(fill = percent_lowage_w,
                          tooltip = toolinfo, data_id = GEOID), 
                      color = "white") +
  scale_fill_gradient(low = "grey90", high = "darkgreen", 
                      name = "Percent\nof Workers\nin Low-wage\nJobs") +
  theme_void() +
  theme(legend.title = element_text(size = 6),
        legend.text = element_text(size = 5),
        legend.position = c(.8, .4))

# make points
gg_point <- ggplot(bg_df, aes(x = percent_lowage_w, 
                              y = fct_rev(county),
                              tooltip = toolinfo, 
                              data_id = GEOID)) +
  geom_point_interactive(size = 7, alpha = 0.2, color = "#674c47") +
  scale_x_continuous(limits = c(0, 100),
                     labels = scales::percent_format(scale = 1),
                     name = "Percent of Population") +
  scale_y_discrete(name = "",
                   labels = function(groups) str_wrap(groups, width = 10)) +
  theme_minimal() +
  theme(legend.title=element_blank(),
        legend.text = element_text(size = 12),
        axis.text.x = element_text(size = 11),
        strip.background = element_rect(color = "black", fill = "#f1f1f1"),
        strip.text.x = element_text(size = 12),
        strip.placement = "outside")

girafe(ggobj = gg_point + gg_map + plot_layout(widths = c(0.45, 0.65)), 
       width_svg = 12, height_svg = 6) %>%
  girafe_options(opts_hover(css = "fill:darkorange;stroke:darkgreen;stroke-width:2px;"))
```

### Residents aged 65 and over
```{r pop2}
# change hover info
bg_df <- bg_df %>% 
  mutate(toolinfo = paste0(names, "\n", 
                           "Percent: ",
                           round(age65per_est, 1), "%",
                           "\n", "Number: ", age65pop_est))

# make map
gg_map <- ggplot(bg_df) +
  # annotation_map_tile(type = "cartolight") +
  geom_sf_interactive(aes(fill = age65per_est,
                          tooltip = toolinfo, data_id = GEOID), 
                      color = "white") +
  scale_fill_gradient(low = "grey90", high = "darkgreen", 
                      name = "Percent of\nResidents\nAged 65\nand over") +
  theme_void() +
  theme(legend.title = element_text(size = 6),
        legend.text = element_text(size = 5),
        legend.position = c(.8, .4))

# make points
gg_point <- ggplot(bg_df, aes(x = age65per_est, 
                              y = fct_rev(county),
                              tooltip = toolinfo, 
                              data_id = GEOID)) +
  geom_point_interactive(size = 7, alpha = 0.2, color = "#674c47") +
  scale_x_continuous(limits = c(0, 100),
                     labels = scales::percent_format(scale = 1),
                     name = "Percent of Population") +
  scale_y_discrete(name = "",
                   labels = function(groups) str_wrap(groups, width = 10)) +
  theme_minimal() +
  theme(legend.title=element_blank(),
        legend.text = element_text(size = 12),
        axis.text.x = element_text(size = 11),
        strip.background = element_rect(color = "black", fill = "#f1f1f1"),
        strip.text.x = element_text(size = 12),
        strip.placement = "outside")

girafe(ggobj = gg_point + gg_map + plot_layout(widths = c(0.45, 0.65)), 
       width_svg = 12, height_svg = 6) %>%
  girafe_options(opts_hover(css = "fill:darkorange;stroke:darkgreen;stroke-width:2px;"))
```

### Residential buildings built before 1970
```{r pop3}
# change hover info
bg_df <- bg_df %>% 
  mutate(toolinfo = paste0(names, "\n", 
                           "Percent: ",
                           round(bldgage70per_est, 1), "%",
                           "\n", "Number: ", bldgage70_est))

# make map
gg_map <- ggplot(bg_df) +
  # annotation_map_tile(type = "cartolight") +
  geom_sf_interactive(aes(fill = bldgage70per_est,
                          tooltip = toolinfo, data_id = GEOID), 
                      color = "white") +
  scale_fill_gradient(low = "grey90", high = "darkgreen", 
                      name = "Percent of\nBuildings\nBuilt\nBefore 1970") +
  theme_void() +
  theme(legend.title = element_text(size = 6),
        legend.text = element_text(size = 5),
        legend.position = c(.8, .4))

# make points
gg_point <- ggplot(bg_df, aes(x = bldgage70per_est, 
                              y = fct_rev(county),
                              tooltip = toolinfo, 
                              data_id = GEOID)) +
  geom_point_interactive(size = 7, alpha = 0.2, color = "#674c47") +
  scale_x_continuous(limits = c(0, 100),
                     labels = scales::percent_format(scale = 1),
                     name = "Percent of Residential Buildings") +
  scale_y_discrete(name = "",
                   labels = function(groups) str_wrap(groups, width = 10)) +
  theme_minimal() +
  theme(legend.title=element_blank(),
        legend.text = element_text(size = 12),
        axis.text.x = element_text(size = 11),
        strip.background = element_rect(color = "black", fill = "#f1f1f1"),
        strip.text.x = element_text(size = 12),
        strip.placement = "outside")

girafe(ggobj = gg_point + gg_map + plot_layout(widths = c(0.45, 0.65)), 
       width_svg = 12, height_svg = 6) %>%
  girafe_options(opts_hover(css = "fill:darkorange;stroke:darkgreen;stroke-width:2px;"))
```

### Households with broadband access and computer
```{r pop4}
# change hover info
bg_df <- bg_df %>% 
  mutate(toolinfo = paste0(names, "\n", 
                           "Percent: ",
                           round(onlineper_est, 1), "%",
                           "\n", "Number: ", onlinehh_est))

# make map
gg_map <- ggplot(bg_df) +
  # annotation_map_tile(type = "cartolight") +
  geom_sf_interactive(aes(fill = onlineper_est,
                          tooltip = toolinfo, data_id = GEOID), 
                      color = "white") +
  scale_fill_gradient(low = "grey90", high = "darkgreen", 
                      name = "Percent of\nHouseholds\nwith Broadband") +
  theme_void() +
  theme(legend.title = element_text(size = 6),
        legend.text = element_text(size = 5),
        legend.position = c(.8, .4))

# make points
gg_point <- ggplot(bg_df, aes(x = onlineper_est, 
                              y = fct_rev(county),
                              tooltip = toolinfo, 
                              data_id = GEOID)) +
  geom_point_interactive(size = 7, alpha = 0.2, color = "#674c47") +
  scale_x_continuous(limits = c(0, 100),
                     labels = scales::percent_format(scale = 1),
                     name = "Percent of Households") +
  scale_y_discrete(name = "",
                   labels = function(groups) str_wrap(groups, width = 10)) +
  theme_minimal() +
  theme(legend.title=element_blank(),
        legend.text = element_text(size = 12),
        axis.text.x = element_text(size = 11),
        strip.background = element_rect(color = "black", fill = "#f1f1f1"),
        strip.text.x = element_text(size = 12),
        strip.placement = "outside")

girafe(ggobj = gg_point + gg_map + plot_layout(widths = c(0.45, 0.65)), 
       width_svg = 12, height_svg = 6) %>%
  girafe_options(opts_hover(css = "fill:darkorange;stroke:darkgreen;stroke-width:2px;"))
```

### Median household income
```{r pop5}
# change hover info
bg_df <- bg_df %>% 
  mutate(toolinfo = paste0(names, "\n", 
                           "Median Income: ", "$",
                           round(medhhinc_est, 1)))

# make map
gg_map <- ggplot(bg_df) +
  # annotation_map_tile(type = "cartolight") +
  geom_sf_interactive(aes(fill = medhhinc_est,
                          tooltip = toolinfo, data_id = GEOID), 
                      color = "white") +
  scale_fill_gradient(low = "grey90", high = "darkgreen", 
                      name = "Median\nHousehold\nIncome") +
  theme_void() +
  theme(legend.title = element_text(size = 6),
        legend.text = element_text(size = 5),
        legend.position = c(.8, .4))

# make points
gg_point <- ggplot(bg_df, aes(x = medhhinc_est, 
                              y = fct_rev(county),
                              tooltip = toolinfo, 
                              data_id = GEOID)) +
  geom_point_interactive(size = 7, alpha = 0.2, color = "#674c47") +
  scale_x_continuous(#limits = c(0, 100),
                     #labels = scales::percent_format(scale = 1),
                     name = "Median Income") +
  scale_y_discrete(name = "",
                   labels = function(groups) str_wrap(groups, width = 10)) +
  theme_minimal() +
  theme(legend.title=element_blank(),
        legend.text = element_text(size = 12),
        axis.text.x = element_text(size = 11),
        strip.background = element_rect(color = "black", fill = "#f1f1f1"),
        strip.text.x = element_text(size = 12),
        strip.placement = "outside")

girafe(ggobj = gg_point + gg_map + plot_layout(widths = c(0.45, 0.65)), 
       width_svg = 12, height_svg = 6) %>%
  girafe_options(opts_hover(css = "fill:darkorange;stroke:darkgreen;stroke-width:2px;"))
```

### Racial identification of residents: Black or African American
```{r pop6}
# change hover info
bg_df <- bg_df %>% 
  mutate(toolinfo = paste0(names, "\n", 
                           "Percent: ",
                           round(blackper_est, 1), "%",
                           "\n", "Number: ", blackpop_est))

# make map
gg_map <- ggplot(bg_df) +
  # annotation_map_tile(type = "cartolight") +
  geom_sf_interactive(aes(fill = blackper_est,
                          tooltip = toolinfo, data_id = GEOID), 
                      color = "white") +
  scale_fill_gradient(low = "grey90", high = "darkgreen", 
                      name = "Percent of\nResidents\nWho Identify\nas Black") +
  theme_void() +
  theme(legend.title = element_text(size = 6),
        legend.text = element_text(size = 5),
        legend.position = c(.8, .4))

# make points
gg_point <- ggplot(bg_df, aes(x = blackper_est, 
                              y = fct_rev(county),
                              tooltip = toolinfo, 
                              data_id = GEOID)) +
  geom_point_interactive(size = 7, alpha = 0.2, color = "#674c47") +
  scale_x_continuous(limits = c(0, 100),
                     labels = scales::percent_format(scale = 1),
                     name = "Percent of Residents") +
  scale_y_discrete(name = "",
                   labels = function(groups) str_wrap(groups, width = 10)) +
  theme_minimal() +
  theme(legend.title=element_blank(),
        legend.text = element_text(size = 12),
        axis.text.x = element_text(size = 11),
        strip.background = element_rect(color = "black", fill = "#f1f1f1"),
        strip.text.x = element_text(size = 12),
        strip.placement = "outside")

girafe(ggobj = gg_point + gg_map + plot_layout(widths = c(0.45, 0.65)), 
       width_svg = 12, height_svg = 6) %>%
  girafe_options(opts_hover(css = "fill:darkorange;stroke:darkgreen;stroke-width:2px;"))
```

### Racial identification of residents: White
```{r pop7}
# change hover info
bg_df <- bg_df %>% 
  mutate(toolinfo = paste0(names, "\n", 
                           "Percent: ",
                           round(whiteper_est, 1), "%",
                           "\n", "Number: ", whitepop_est))

# make map
gg_map <- ggplot(bg_df) +
  # annotation_map_tile(type = "cartolight") +
  geom_sf_interactive(aes(fill = whiteper_est,
                          tooltip = toolinfo, data_id = GEOID), 
                      color = "white") +
  scale_fill_gradient(low = "grey90", high = "darkgreen", 
                      name = "Percent of\nResidents\nWho Identify\nas White") +
  theme_void() +
  theme(legend.title = element_text(size = 6),
        legend.text = element_text(size = 5),
        legend.position = c(.8, .4))

# make points
gg_point <- ggplot(bg_df, aes(x = whiteper_est, 
                              y = fct_rev(county),
                              tooltip = toolinfo, 
                              data_id = GEOID)) +
  geom_point_interactive(size = 7, alpha = 0.2, color = "#674c47") +
  scale_x_continuous(limits = c(0, 100),
                     labels = scales::percent_format(scale = 1),
                     name = "Percent of Residents") +
  scale_y_discrete(name = "",
                   labels = function(groups) str_wrap(groups, width = 10)) +
  theme_minimal() +
  theme(legend.title=element_blank(),
        legend.text = element_text(size = 12),
        axis.text.x = element_text(size = 11),
        strip.background = element_rect(color = "black", fill = "#f1f1f1"),
        strip.text.x = element_text(size = 12),
        strip.placement = "outside")

girafe(ggobj = gg_point + gg_map + plot_layout(widths = c(0.45, 0.65)), 
       width_svg = 12, height_svg = 6) %>%
  girafe_options(opts_hover(css = "fill:darkorange;stroke:darkgreen;stroke-width:2px;"))
```

### Ethnic identification of residents: Hispanic or Latino
```{r pop8}
# change hover info
bg_df <- bg_df %>% 
  mutate(toolinfo = paste0(names, "\n", 
                           "Percent: ",
                           round(ltnxper_est, 1), "%",
                           "\n", "Number: ", ltnxpop_est))

# make map
gg_map <- ggplot(bg_df) +
  # annotation_map_tile(type = "cartolight") +
  geom_sf_interactive(aes(fill = ltnxper_est,
                          tooltip = toolinfo, data_id = GEOID), 
                      color = "white") +
  scale_fill_gradient(low = "grey90", high = "darkgreen", 
                      name = "Percent of\nResidents\nWho Identify\nas Hispanic") +
  theme_void() +
  theme(legend.title = element_text(size = 6),
        legend.text = element_text(size = 5),
        legend.position = c(.8, .4))

# make points
gg_point <- ggplot(bg_df, aes(x = ltnxper_est, 
                              y = fct_rev(county),
                              tooltip = toolinfo, 
                              data_id = GEOID)) +
  geom_point_interactive(size = 7, alpha = 0.2, color = "#674c47") +
  scale_x_continuous(limits = c(0, 100),
                     labels = scales::percent_format(scale = 1),
                     name = "Percent of Residents") +
  scale_y_discrete(name = "",
                   labels = function(groups) str_wrap(groups, width = 10)) +
  theme_minimal() +
  theme(legend.title=element_blank(),
        legend.text = element_text(size = 12),
        axis.text.x = element_text(size = 11),
        strip.background = element_rect(color = "black", fill = "#f1f1f1"),
        strip.text.x = element_text(size = 12),
        strip.placement = "outside")

girafe(ggobj = gg_point + gg_map + plot_layout(widths = c(0.45, 0.65)), 
       width_svg = 12, height_svg = 6) %>%
  girafe_options(opts_hover(css = "fill:darkorange;stroke:darkgreen;stroke-width:2px;"))
```