---
title: "Place Names"
date: "6/23/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tidyverse)
library(leaflet)
library(ggthemes)

tracts <- readRDS("tractnames.RDS")
pal <- colorFactor(palette = tableau_color_pal(palette = "Tableau 20")(17), tracts$TRACTCE)

```

## Eastern Shore Place Names {.tabset}
The maps below summarize the notes and names generated from the Community Advisory Committee.

* Move between the tabs to see the tracts overlaid on different base maps with different underlying details.
* Zoom and center a map to view the area defined by a census tract more closely.
* Click anywhere within a tract to review the notes and potential names for that tract.

### Base OSM
```{r map2, fig.fullwidth = TRUE}
leaflet(tracts) %>% 
  addTiles() %>% 
  #addProviderTiles("Esri.NatGeoWorldMap") %>% 
  addPolygons(fillColor = ~ pal(TRACTCE),
              fillOpacity = 0.3,
              color = ~ pal(TRACTCE),
              weight = 2,
              popup = paste0("Tract Number: ", tracts$TRACTCE, "<br>",
                             "Notes/Names: ", tracts$names))
```

### Base Esri
```{r map1, fig.fullwidth = TRUE}
leaflet(tracts) %>% 
  #addTiles() %>% 
  addProviderTiles("Esri.DeLorme") %>% 
  addPolygons(fillColor = ~ pal(TRACTCE),
              fillOpacity = 0.3,
              color = ~ pal(TRACTCE),
              weight = 2,
              popup = paste0("Tract Number: ", tracts$TRACTCE, "<br>",
                             "Notes/Names: ", tracts$names))
```


